---
title: "HW2"
author: "Alaia Rubio"
date: "2025-02-27"
output: html_document
---

```{r}
library(ggplot2)
```


Q1. Using mpg dataset, create a scatter plot between displ (Displacement) and hwy (Highway mileage)

```{r basic_plot}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point()

```

Question 2 on Color and Shape of datapoints 
- Create scatter plot between displ (on x-axis) and hwy (y-axis), to include different shapes for the datapoints by drv (Drivetrain type) and different color by the class (class type)



```{r geom_additions}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, shape = drv, color = class)) +
  geom_point()

```

  

Q3. Create a boxplot on the "hwy" mileage. Make sure to use the position argument in the geom_ function and the "outlier.colour" property. The outlier is highlighted in "red" color. 
- The default statistics of geom_boxplot is statistics. It is embedded in the stat parameter.

```{r boxplot1}

ggplot(data = mpg, aes(x = hwy)) +
  geom_boxplot(outlier.colour = "red",
               outlier.shape = 16,
               outlier.size = 2.5,
               position = position_identity()) +
  labs(title = "Boxplot of Highway Mileage (hwy)",
       x= "Highway Mileage (mpg)",
       y= "") +
  theme_minimal()

```
- Q4. Add labels to the bar graph as shown below. The labels represent the count of specific class cars in the mpg dataset. You could add labels using geom_text() function. Therefore, one way to recreate the bar graph with labels is to stack, geom_bar followed by geom_text. 
- Hint: stat = "count" statistic for calculating the count of each class, should be used in the geom_text in case if you are trying to add the labels using the geom_text.

```{r stacking_geoms}
ggplot(data = mpg, mapping = aes(x = class)) +
  geom_bar(fill = "darkgrey") +
  geom_text(stat = "count", aes(label = after_stat(count)), vjust = -0.5) +
  labs(x = "Car Class", y = "Count")

```
- Q5. Use scale_x_continuous to add a label to the x-axis. Aim for a closest possible replica of the provided graph.

```{r scales1}

ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = class, shape = drv)) +
  geom_point() +
  scale_x_continuous(name = 'Using "hwy" variable for Gas Mileage') +
  labs(y = "hwy")


```
- Q6. Use lims function to limit the x variable displ to display only the vehicles between displ value of 1 to 6, and hwy vehicles with mileage between 10 and 40. Aim for a closest possible replica of the provided graph.

```{r scales_limit}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, 
                                 color = class, shape = drv)) +
  geom_point() +
  lims(x = c(1, 6), y = c(10, 40)) +
  labs(x = "displ", y = "hwy")

```
- Q7. Use facet_wrap or fact_grid on manufacturer variable to display scatter plot between variables x = displ and y = hwy. Aim for a closest possible replica of the provided graph.

```{r facets1}


ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = class, shape = drv)) +
  geom_point() +
  facet_wrap(~ manufacturer, ncol = 4) +
  labs(x = "displ", y = "hwy")

```
- Q8. Create bar graph by class. Experiment with linetype argument and give a different outline color using color argument.

```{r coordinate1-base}

ggplot(data = mpg, mapping = aes(x = class, fill = drv)) +
  geom_bar(color = "black", linetype = "dashed", position = position_dodge()) +
  labs(x = "class", y = "COUNT")

```
- Q9. Use coord_flip on the graph from Question 8, Q8. and flip the coordinates 
- Flip the coordinate 

```{r}
ggplot(data = mpg, mapping = aes(x = class, fill = drv)) +
  geom_bar(color = "black", position = position_dodge()) +
  labs(x = "class", y = "count") +
  coord_flip()

```
- Q10. Use polar coordinates to create the graph.
- Display the same graph on a polar coordinate system 

```{r coordinate2}


ggplot(data = mpg, mapping = aes(x = class, fill = drv)) +
  geom_bar(color = "black", position = position_dodge()) +
  labs(x = "class", y = "count") +
  coord_polar("x") 

```